Applied Measure Theory for Probabilistic Modeling

### 2022-05-15 paper/paper.tex
edits

Given a \emph{parameter space} $\Theta$ and \emph{observation space} $X$, a \verb|Likelihood| $\ell$ consists of a kernel $\ell_κ : \Theta \to \cM(X)$ and an observation $\ell_x \in X$. Given a measure $\rho \in \cM(X)$, $\ell$ can be treated as a function $\Theta \to \RR_{\ge 0}$, given by
\[
    \ell(\theta) \equiv \frac{\dd\ell_\kappa(\theta)}{\dd\rho}(\ell_x)
\]

In \verb|MeasureTheory|, we take $\rho$ to be the \emph{root measure} of $\ell_\kappa(\theta)$
Note that $\ell$ takes a point in $\Theta$ to an evaluated density in a different space (the observation space $X$). In particular, a likelihood is neither a measure nor a density.

>>>>

Given a \emph{parameter space} $\Theta$ and \emph{observation space} $X$, a \verb|Likelihood| $\ell$ consists of 
\begin{enumerate}
    \item a kernel $\ell_κ : \Theta \to \cM(X)$,
    \item an observation $\ell_x \in X$, and
    \item a base measure $\ell_\beta \in\cM(X)$.
\end{enumerate}

A likelihood can be treated as a function $\Theta \to \RR_{\ge 0}$, by defining
\[
    \ell(\theta) = \frac{\dd\ell_\kappa(\theta)}{\dd\ell_\beta}(\ell_x)
\]

Of particular interest for the base measure is the case
\[
    \ell_\beta = \hat{\theta} \equiv \arg\max_\theta \frac{\dd\ell_\kappa(\theta)}{\dd\rho}(\ell_x)
\]

for any non-singular measure $\rho$; this corresponds to the \href{https://en.wikipedia.org/wiki/Likelihood_function#Relative_likelihood_function}{\emph{relative likelihood}}. In \verb|MeasureTheory|, we allow the base measure to be specified, but take as a default the \emph{root measure} of $\ell_\kappa(\theta)$.

Note that a likelihood $\ell$ takes a point in $\Theta$ to an evaluated density in a different space (the observation space $X$). In particular, a likelihood is neither a measure nor a density.
